Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-02-03 article/article.Rmd
* fix readmes * add in surrogacy tests * progress on article * added in figure in figure.Rmd

Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning. First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritizations. Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritizations. Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritization and the proportion of genetic variation secured in a prioritization. If these surrogates can be shown to markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

### Landscape data
We aimed to investigate the effectiveness of climatic and geographic surrogates. To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates. To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). We used the total human population size inside each planning unit to represent its opportunity cost (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @rXXX).

@@

To understand how prioritizations generated using surrogate-based targets might perform in conservation planning scenarios, we generated three sets of prioritizations: a set of single-species prioritizations, a set of multi-species prioritizations, and a set of multi-species prioritizations that accommodated opportunity cost. For each of these sets, we generated prioritizations using only amount-based targets, prioritizations using both amount- and surrogate-based targets, and prioritizations using amount- and genetic-based targets. All amount-based targets were set as `r (rapr.params.LST[[MODE]]$amount.target*100)` % to ensure an adequate proportion of habitat was secured for each species [@rXXX]. All surrogate- and genetic-based targets were set as `r (rapr.params.LST[[MODE]]$surrogate.target*100)`. For each set, for each combination of targets, we identified the `r rapr.params.LST[[MODE]]$single.species.surrogate.replicates` solutions nearest to optimality for replication. Except for the single-species amount-only prioritizations for which we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` replicates because they were associated with many optimal solutions with varying levels of effectiveness.

To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated two sets of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %. The first set of prioritizations contained prioritizations with varying environmental surrogate-based targets with no specified geographic targets. The second contained prioritizations with varying geographic surrogate-based targets and no environmental targets. None of these prioritizations used amount-based targets.

## Statistical analysis
We tested for differences in the effectiveness of prioritizations representing typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469]. We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link. Both models contained a predictor variable indicating if the prioritization was generated using amount-only, surrogate, or genetic targets, a second predictor variable indicating if the genetic variation was adaptive or neutral, and the interaction between these variables. Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species. Each full model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package). Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate). Both models accommodated species-level random effects for the asymptote ($a$) and rate ($r$) parameters.

>>>>

Our overarching aim was to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning. First, we aimed to understand the relationship between the surrogate targets used to generate a prioritization and the proportion of genetic variation secured in the prioritisation. Second, we aimed to determine if the addition of surrogate-based targets in conservation planning exercises would result in prioritizations that secured substantially more genetic variation. If these surrogatess can be shown to improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

### Landscape data
We aimed to investigate the effectiveness of climatic and geographic surrogates. To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates. To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @r472).

@@

To understand the relationship between the surrogate-based targets used to generate a prioritizations and the amount of genetic variation secured, we generated a collection of single-species prioritizations. For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %. None of these prioritizations contained any amount-based targets.

To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we used three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. Most large-scale prioritizations involve a comprehensive set of species, and in turn, may secure a greater proportion of intra-specific genetic variation. Our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations. We generated prioritizations using only amount-based, both amount- and surrogate-based targets, and both amount- and genetic-based targets. The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline. We used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX]. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % to ensure that most of the repsective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.parmas.LST[[MODE]]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with a single optimal solution, we used only a single replicate.

## Statistical analysis
We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMMs; using the $\texttt{lme4}$ R package; @r469]. A model was fit to the genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate). An interaction term was included to estimate parameters separately for adaptive and neutral variation. The models accommodated species-level variation using random effects for both parameters.

We tested for differences in the effectiveness of prioritizations generated under different scenarios and targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package]. We fit a full model to the proportion of genomic variation secured in all prioritizations using a logit link. Both models contained variables indicating the scenario used to generate the prioritization, the targets used to generate the prioritizations, and the type of genetic variation measured. The model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). All models contained an intercept-only random effect to ensure that comparisons were made within species. Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

### 2016-02-03 article/figures.Rmd
* fix readmes * add in surrogacy tests * progress on article * added in figure in figure.Rmd

>>>>

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogates and adaptive genetic variation. Panel (b) shows that for geographic surrogates and neutral genetic variation. Transparent gray lines denote trends for each species, and black lines show general trends across all species.'}
### prepare data frames for plotting
## environmental vs adaptive
# random effects
env.correlation.RPDF <- expand.grid(
	Surrogate.type='Environmental variation',
	Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(env.correlation.DF$Species)
)
env.correlation.RPDF$a.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$a[env.correlation.RPDF$Species]
env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[2]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
env.correlation.RPDF$fit <- env.correlation.RPDF$a.ranef - (env.correlation.RPDF$a.ranef - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
# fixed effects
env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
env.correlation.FPDF$fit <- fixef(env.correlation.NLMM)[[1]] - (fixef(env.correlation.NLMM)[[1]] - 0) * exp(- fixef(env.correlation.NLMM)[[2]] * env.correlation.FPDF$Surrogate.target)
## geographic vs neutral
# random effects
geo.correlation.RPDF <- expand.grid(
	Surrogate.type='Geographic variation',
	Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(geo.correlation.DF$Species)
)
geo.correlation.RPDF$a.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$a[geo.correlation.RPDF$Species]
geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[2]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
geo.correlation.RPDF$fit <- geo.correlation.RPDF$a.ranef - (geo.correlation.RPDF$a.ranef - 0) * exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target)
# fixed effects
geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
geo.correlation.FPDF$fit <- fixef(geo.correlation.NLMM)[[1]] - (fixef(geo.correlation.NLMM)[[1]] - 0) * exp(- fixef(geo.correlation.NLMM)[[2]] * geo.correlation.FPDF$Surrogate.target)

### make plots
suppressWarnings({
	p1 <- ggplot() +
		geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
		geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
		geom_text(aes(x=0.6, y=1.03, label='(a)')) +
		xlab('Environmental variation secured (%)') +
		ylab('Adaptive genetic\nvariation secured (%)') +
		theme_classic()
	p2 <- ggplot() +
		geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
		geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
		geom_text(aes(x=0.6, y=1.03, label='(b)')) +
		xlab('Geographic variation secured (%)') +
		ylab('Neutral genetic\nvariation secured (%)') +
		theme_classic()
	grid.arrange(p1, p2, nrow=1)
})
```
